---
title: "hydro_climate_trends"
author: "Nick Gubbins"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages.
library(here)
source(here('src', 'setup.R'))

```
# Trend detection

## Reading in and summarizing data.

First reading in summary data and work that Heili already did.

```{r}
# if you want to save runtime comment out source(q_metrics.r)
#source(here('src', 'q_metrics.R'))
com <- readRDS(here('data_working', 'discharge_metrics_siteyear.rds')) %>%
    select(-contains('date')) %>% #removing redundant date columns that are harder to trend
    filter(water_year >= landsat_year)%>%
    reduce_to_longest_site_runs()

com_long <- pivot_longer(com, cols = -c('site_code', 'water_year'), names_to = 'var', values_to = 'val')
```

Perform trend detection at each site.

```{r}
com_trends <- detect_trends(com_long, 'longest_site_run_landsat') %>%
    add_flags()
write_csv(com_trends, here('data_working','trends','longest_site_run_landsat.csv'))
#com_trends <- read_csv(here('data_working', 'hydro_climate_trends.csv'))
```

# Figure creation

## Trend histograms/density plots

First want an idea of what data is in here. So going to make a quick table.

```{r}
com %>%
    left_join(., ms_site_data, by = 'site_code') %>%
    select(domain, site_code) %>%
    group_by(domain) %>%
    summarize(n = n_distinct(site_code))
```

```{r}
full<- as.integer(length(unique(com_trends$var)))
half <- as.integer(length(unique(com_trends$var))/2 )
```

Make kernel density plots for each trend. 

```{r}
com_trends %>%
    filter(var %in% unique(com_trends$var)[1:half]) %>%
ggplot(., aes(x = trend, fill = flag))+
    geom_density()+
    facet_wrap(~var, scales = 'free')+
    theme_few()+
    labs(title = 'Trends')+
    scale_fill_manual(values = flag_colors)
```



```{r}
com_trends %>%
    filter(var %in% unique(com_trends$var)[half+1:full]) %>%
ggplot(., aes(x = trend, fill = flag))+
    geom_density()+
    facet_wrap(~var, scales = 'free')+
    theme_few()+
    labs(title = 'Trends')+
    scale_fill_manual(values = flag_colors)
```


```{r}
com_trends %>%
    filter(var %in% c('m1_meanq', 'q1', 'q50', 'q99')) %>%
    ggplot(., aes(x = trend, fill = flag))+
    geom_density()+
    facet_wrap(~var, scales = 'free')+
    theme_few()+
    labs(title = 'Trends')+
    scale_fill_manual(values = flag_colors)
```


```{r}
com_trends %>%
    filter(var %in% unique(com_trends$var)[1:half]) %>%
ggplot(.,aes(x = trend, fill = flag))+
    geom_histogram(position = 'dodge')+
    facet_wrap(~var, scales = 'free', ncol = 2)+
    theme_few()+
    labs(title = 'Trends')+
    scale_fill_manual(values = flag_colors)
```


```{r}
com_trends %>%
    filter(var %in% unique(com_trends$var)[half+1:full])%>%
ggplot(.,aes(x = trend, fill = flag))+
    geom_histogram(position = 'dodge')+
    facet_wrap(~var, scales = 'free', ncol = 2)+
    theme_few()+
    labs(title = 'Trends')+
    scale_fill_manual(values = flag_colors)
```



```{r}
com_trends %>%
    filter(var %in% c('m1_meanq', 'q1', 'q50', 'q99')) %>%
    ggplot(., aes(x = trend, fill = flag))+
    geom_histogram(position = 'dodge')+
    facet_wrap(~var, scales = 'free', ncol = 2)+
    theme_few()+
    labs(title = 'Trends')+
    scale_fill_manual(values = flag_colors)
```

Make kernal density plot for each trend, but at the variable level. Split out trends as facets. Display data as lines in accessory panel for context. See issue #16 on github for example.

Wow this is tricky. Going to try adapting their code from the paper here. 

https://github.com/bellaoleksy/rocky-mtn-color/blob/main/3_Trends_In_Color.Rmd


## Scatters

Make scatter plots to put domains in context. Going to pull average each domain for now, will want to pick specific exemplar sites in the near future.


### Prep data
```{r}
#uncomment this if using the v2 dataset from the ms package proper
#ms_site_data <- ms_load_sites()

context_site <- com_long %>%
    left_join(com_trends, by = c('site_code', 'var')) %>%
    left_join(., ms_site_data, by = 'site_code') %>%
    filter(water_year %in% 2000:2019,
           site_code %in% unique(trends$site_code)) #only looking at data we have trends for


context_domain <- com %>%
    left_join(., ms_site_data, by = 'site_code') %>%
    filter(water_year %in% 2000:2019,
           site_code %in% unique(trends$site_code)) %>% #only looking at data we have trends for
    mutate(domain = case_when(domain == 'neon' ~ paste0(domain, '_', site_code),
                              domain != 'neon' ~ domain)) %>% # disaggregate neon sites since its national scale
    group_by(domain) %>%
    summarize(flow = mean(m1_meanq, na.rm = T),
              temp = mean(temp_mean_ann, na.rm = T),
              precip = mean(precip_mean_ann, na.rm = T)) %>%
    mutate(runoff_ratio = flow/precip)
```

### Make generic plotter to make context plots easier on us.

```{r}
# first make domain level
plot_context_domain <- function(x_in = NA, y_in = NA, target_trend = NA){
  
x_in <- 'temp'
y_in <- 'precip' 


ggplot(context_domain, aes_string(x = x_in, y = y_in, color = 'domain')) +
    geom_point()+
    geom_text_repel(aes(label = domain),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 30)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)    
}
```

Next make site level plotter

```{r}
plot_context_site <- function(x_in = NA, y_in = NA, target_trend = NA){

x_in <- x_in
y_in <- y_in
target_trend <- target_trend

start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    filter(var %in% c(x_in, y_in)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val') %>%
    na.omit()

trend_fil <- com_trends %>%
    filter(var == target_trend)

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code')

    p_all <- ggplot(plot_dat, aes_string(x = x_in, y = y_in, color = 'domain'))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        scale_color_viridis(discrete = T)+
        labs(title = 'All')
 
# good chance this will break if they update ggplot!   
 x_breaks <- ggplot_build(p_all)$layout$panel_params[[1]]$x.sec$breaks
 y_breaks <- ggplot_build(p_all)$layout$panel_params[[1]]$y.sec$breaks
 x_lims <- ggplot_build(p_all)$layout$panel_params[[1]]$x.sec$limits
 y_lims <- ggplot_build(p_all)$layout$panel_params[[1]]$y.sec$limits
 
    p_up <- plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(aes_string(x = x_in, y = y_in, color = 'domain'))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        scale_color_viridis(discrete = T)+
        labs(title = paste0('Increasing ', target_trend))+
        scale_x_continuous(breaks = x_breaks,
                           limits = x_lims)+
        scale_y_continuous(breaks = y_breaks,
                           limits = y_lims)
    
    p_down <- plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(aes_string(x = x_in, y = y_in, color = 'domain'))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        scale_color_viridis(discrete = T)+
        labs(title = paste0('Decreasing ', target_trend))+
        scale_x_continuous(breaks = x_breaks,
                           limits = x_lims)+
        scale_y_continuous(breaks = y_breaks,
                           limits = y_lims)
    
    p_non <- plot_dat %>%
        filter(flag == 'non-significant') %>%
    ggplot(aes_string(x = x_in, y = y_in, color = 'domain'))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        scale_color_viridis(discrete = T)+
        labs(title = paste0('No dectectable trend in ', target_trend))+
        scale_x_continuous(breaks = x_breaks,
                           limits = x_lims)+
        scale_y_continuous(breaks = y_breaks,
                           limits = y_lims)
       
p_out <- (p_all + p_non) / (p_up + p_down)
p_out
}

```

### Apply generic plotter

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'rbiq')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'm1_meanq')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'm5_ar1q')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'runoff_ratio')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'precip_mean_ann')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'precip_mean_ann')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'mintemp_winter')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'meantemp_summer')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'q50_dowy_exceed')
```

```{r}
plot_context_site(x_in = 'precip_mean_ann', y_in = 'runoff_ratio', target_trend = 'p50_dowy_exceed')
```


### precip & temp

```{r}
ggplot(context_domain, aes(x = temp, y = precip, color = domain)) +
    geom_point()+
    geom_text_repel(aes(label = domain),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 30)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)
```


Break out plots for m1, ar1, rbi.

First m1:

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'm1_meanq') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code')

    ggplot(plot_dat, aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'All')
    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing Flow')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing Flow')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
```

Ar1:

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'm5_ar1q') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code')
    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing AR1')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing AR1')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
```


Rbi

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'rbiq') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code')
    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing RBI')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing RBI')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
```

temp

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'temp_mean_ann') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code')
    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing Temp')+
        scale_x_continuous(breaks = c(0,5,10,15))+
        scale_y_continuous(breaks = c(2,4,6))
    
```

precip

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'precip_mean_ann') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code')
    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing Precip')+
        scale_x_continuous(breaks = c(0,5,10,15),
                           limits = c(0,15))+
        scale_y_continuous(breaks = c(2,4,6),
                           limits = c(0,6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(aes(x = temp_mean_ann, y = precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'T (mean annual, C)', y = 'Precip (mean annual, mm/d)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing Precip')+
        scale_x_continuous(breaks = c(0,5,10,15),
                           limits = c(0,15))+
        scale_y_continuous(breaks = c(2,4,6),
                           limits = c(0,6))
    
    
```

### Flow & Precip

```{r}
context_domain %>%
    filter(domain != 'neon_MCDI') %>%
ggplot(., aes(x = precip, y = flow, color = domain)) +
    geom_point()+
    geom_text_repel(aes(label = domain),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 30,
                  force = 20)+
        geom_abline(slope = 1, color = 'red')+
        theme_few(base_size = 20) +
        theme(legend.position = 'none')+
        labs(x = 'Precip (mean annual, mm/d)', y = 'Q (mean annual, mm/d)',
             caption = 'Note: NEON_MCDI off scale Q. (Q 270, P 2.6)')+
        scale_color_viridis(discrete = T) 


```

### runoff ratio to precip (all)

```{r}
context %>%
    filter(domain != 'neon_MCDI') %>%
ggplot(., aes(x = precip, y = runoff_ratio, color = domain)) +
    geom_point()+
    geom_text_repel(aes(label = domain),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 30,
                  force = 20)+
        geom_hline(yintercept= 1, color = 'red')+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)
```

For mean q:

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'm1_meanq') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code') %>%
    filter(site_code != 'MCDI')


    ggplot(plot_dat, aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        scale_y_continuous(limits = c(0, 6))+
        labs(title = 'All')
    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing Flow')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing Flow')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6))
    
```

for ar1

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'm5_ar1q') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code') %>%
    filter(site_code != 'MCDI')

    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing AR1')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing AR1')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
```

rbi:

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'rbiq') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code') %>%
    filter(site_code != 'MCDI')

    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing RBI')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing RBI')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
```

for precip

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'precip_mean_ann') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code') %>%
    filter(site_code != 'MCDI')

    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing Precip')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
    
    plot_dat %>%
        filter(flag == 'decreasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Decreasing Precip')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
```


for temp:

```{r}
start <- context_site %>%
    group_by(site_code, var, domain) %>%
    summarize(val = mean(val, na.rm = T)) %>%
    pivot_wider(id_cols = c('site_code', 'domain'), names_from = 'var', values_from = 'val')

trend_fil <- com_trends %>%
    filter(var == 'temp_mean_ann') 

plot_dat <- start %>%
    left_join(., trend_fil, by = 'site_code') %>%
    filter(site_code != 'MCDI')

    
    plot_dat %>%
        filter(flag == 'increasing') %>%
    ggplot(., aes(x = precip_mean_ann, y = m1_meanq/precip_mean_ann, color = domain))+
        geom_text(aes(label = site_code, size = .0001))+
        geom_mark_ellipse(aes(label = domain), label.fontsize = 20)+
        scale_color_viridis(discrete = T)+
        theme_few(base_size = 20) +
        geom_hline(yintercept= 1, color = 'red')+
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             caption = 'Note: neon_MCDI off scale (RR 87, P 2.3)')+
        scale_color_viridis(discrete = T)+
        labs(title = 'Increasing Temp')+
        scale_y_continuous(limits = c(0, 6))+
        scale_x_continuous(breaks = c(2,4,6),
                           limits = c(0, 6))
    
```



### runoff ratio to precip (-neon)

```{r}
context %>%
    mutate(runoff_ratio = flow/precip) %>%
    filter(!grepl('neon', domain)) %>%
ggplot(., aes(x = precip, y = runoff_ratio, color = domain)) +
    geom_point()+
    geom_text_repel(aes(label = domain),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 30,
                  force = 20)+
        theme_few(base_size = 20) +
        theme(legend.position = 'none') +
        labs(x = 'Precip (mean annual, mm/d)', y = 'Runoff Ratio',
             title = 'neon excluded')+
        scale_color_viridis(discrete = T)
```

## Boxplots

### CAMELS chem comparison
Re-create part of figure 5 from the CAMEL CHEM data paper (G sterle et al 2024).

Need a vertical boxplot of mean daily precip and Q.

```{r}
context_domain %>%
    pivot_longer(cols = -domain, names_to = 'var', values_to = 'val') %>%
    filter(var == 'flow') %>%
    select(var, val) %>%
    ggplot(., aes(y = val))+
    geom_boxplot()+
    theme_few(base_size = 30)+
    labs(title = 'mean daily Q')+
    scale_y_continuous(breaks = c(0, 2.5, 5, 7.5, 10, 12.5)) # set breaks to match their figure

```


```{r}
context_domain %>%
    pivot_longer(cols = -domain, names_to = 'var', values_to = 'val') %>%
    filter(var == 'precip') %>%
    select(var, val) %>%
    ggplot(., aes(y = val))+
    geom_boxplot()+
    theme_few(base_size = 30)+
    labs(title = 'mean daily P')+
    scale_y_continuous(breaks = c(0, 2.5, 5, 7.5, 10)) # set breaks to match their figure

```

